---
title: "Phenology_Models"
output: html_document
date: "2025-01-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse)
library(lubridate)
library(lme4)
library(bbmle)
library(googlesheets4)
library(lmtest)
library(car)
library(lmerTest)
```

```{r}
UCD22_23 <- read_csv("../input/UCD22_23_transplant.csv")

WL2_23 <- read_csv("../input/WL2_23.csv")
WL2_24 <- read_csv("../input/WL2_24.csv")

clim_dist_UCD22_23 <- read_csv("../input/growthseason_GowersEnvtalDist_UCD.csv")
clim_dist_WL2_23 <- read_csv("../input/growthseason_GowersEnvtalDist_WL2.csv")
```

#### Select only the columns of interest and only parent pops from the WL2_24 garden

```{r}
UCD22_23 <- UCD22_23 %>%
  mutate(bed = str_extract(block, "^[A-Za-z]")) %>% #create column for bed by extracting info from block column
  select(block, bed, row, col, everything()) %>% #reorder columns
  select(block, bed, row, col, pop, mf, rep, bud.date_dt, flower.date_dt, fruit.date_dt, last.FL.date_dt, last.FR.date_dt, death.date_dt) %>% #select columns of interest
  filter(pop != "buffer") #filter out buffer plants

WL2_23 <- WL2_23 %>%
  select(block, bed, row, col, pop, mf, rep, bud.date_dt, flower.date_dt, fruit.date_dt, last.FL.date_dt, last.FR.date_dt, death.date_dt) %>%
  filter(grepl("^\\d+$", rep)) #removes buffer plants and empty spaces

WL2_24 <- WL2_24 %>%
  select(bed, row, col, pop, mf, rep, bud.date_dt, flower.date_dt, fruit.date_dt, last.FL.date_dt, last.FR.date_dt, death.date_dt) %>%
  left_join(WL2_23 %>% select(block, bed, row, col), by=c("bed", "row", "col")) %>% #add block info using WL2_23 block column
  select(block, bed, row, col, everything()) #reorder columns 
```

#### Combine all gardens into one tibble, add a site column

```{r}
parent_pheno <- bind_rows(
  UCD22_23 %>% mutate(site="UCD_23"),
  WL2_23 %>% mutate(site="WL2_23"),
  WL2_24 %>% mutate(site="WL2_24")
)

parent_pheno <- parent_pheno %>%
  separate(site, into = c("site", "year"), sep = "_")
```

#### Add elevation data

```{r}
clim_dist_UCD22_23 <- clim_dist_UCD22_23 %>%
  filter(TimePd == "Recent") %>%
  rename(pop=parent.pop)

parent_pheno <- parent_pheno %>%
  left_join(clim_dist_UCD22_23 %>% select(pop, elev_m), by="pop")
```

#### Calculate duration of each reproductive stage and add binary column for initiating reproduction

```{r}
#Bud duration (flower.date - bud.date)
#Flower duration (last.FL.date - flower.date)
#Fruit duration (last.FR.date - fruit.date)
#Total repro duration (last.FR.date - bud.date)
#Initiate repro(1 if there is any repro date, 0 if no repro date)
  # I would want to use bud.date to determine if a plant initiated repro, but we missed the bud date for some plants

parent_pheno <- parent_pheno %>%
  mutate(bud.duration = as.numeric(flower.date_dt - bud.date_dt)) %>%
  mutate(flower.duration = as.numeric(last.FL.date_dt - flower.date_dt)) %>%
  mutate(fruit.duration = as.numeric(last.FR.date_dt - fruit.date_dt)) %>%
  mutate(total.repro.duration = as.numeric(last.FR.date_dt - bud.date_dt)) %>%
  mutate(initiate.repro = if_else(
      !is.na(bud.date_dt) | !is.na(flower.date_dt) | 
      !is.na(fruit.date_dt) | !is.na(last.FR.date_dt) | 
      !is.na(last.FL.date_dt), 
      1, 0
    ))
```

#### Visualizing each repro duration across sites

```{r}
parent_pheno %>%
  filter(!is.na(bud.duration)) %>%
  filter(!(pop %in% c("CP2", "SQ1", "SQ3", "WR"))) %>%
  ggplot(aes(x=bud.duration, fill=site)) +
  geom_boxplot() +
  facet_wrap(~pop)

parent_pheno %>%
  filter(!is.na(flower.duration)) %>%
  #filter(!(pop %in% c("CP2", "SQ1", "SQ3", "WR"))) %>%
  ggplot(aes(x=flower.duration, fill=site)) +
  geom_boxplot() +
  facet_wrap(~pop)

parent_pheno %>%
  filter(!is.na(fruit.duration)) %>%
  #filter(!(pop %in% c("CP2", "SQ1", "SQ3", "WR"))) %>%
  ggplot(aes(x=fruit.duration, fill=site)) +
  geom_boxplot() +
  facet_wrap(~pop)

parent_pheno %>%
  filter(!is.na(total.repro.duration)) %>%
  #filter(!(pop %in% c("CP2", "SQ1", "SQ3", "WR"))) %>%
  ggplot(aes(x=total.repro.duration, fill=site)) +
  geom_boxplot() +
  facet_wrap(~pop)
```

#### Visualize days to each repro stage

```{r}
parent_pheno %>%
  filter(!is.na(bud.date_dt)) %>%
  #filter(!(pop %in% c("CP2", "SQ1", "SQ3", "WR", "YO7", "IH"))) %>%
  ggplot(aes(x=bud.date_dt, fill=site)) +
  geom_boxplot() +
  facet_wrap(~pop) #ALL YO7 budded on the same date  at WL2, All WL2 budded on the same date at WL2 (only 1 WL2 reproduced at UCD), TM2 flowering at UCD is very syncronous

parent_pheno %>%
  filter(!is.na(flower.date_dt)) %>%
  #filter(!(pop %in% c("CP2", "SQ1", "SQ3", "WR"))) %>%
  ggplot(aes(x=flower.date_dt, fill=site)) +
  geom_boxplot() +
  facet_wrap(~pop)

parent_pheno %>%
  filter(!is.na(fruit.date_dt)) %>%
  #filter(!(pop %in% c("CP2", "SQ1", "SQ3", "WR"))) %>%
  ggplot(aes(x=fruit.date_dt, fill=site)) +
  geom_boxplot() +
  facet_wrap(~pop)

parent_pheno %>%
  filter(!is.na(last.FL.date_dt)) %>%
  #filter(!(pop %in% c("CP2", "SQ1", "SQ3", "WR"))) %>%
  ggplot(aes(x=last.FL.date_dt, fill=site)) +
  geom_boxplot() +
  facet_wrap(~pop)

parent_pheno %>%
  filter(!is.na(last.FR.date_dt)) %>%
  #filter(!(pop %in% c("CP2", "SQ1", "SQ3", "WR"))) %>%
  ggplot(aes(x=last.FR.date_dt, fill=site)) +
  geom_boxplot() +
  facet_wrap(~pop)


```

#### Model: Does initiating reproduction differ by site/elevation?

```{r}
#Does initiating reproduction differ by site with pop as a random effect
initiate_site <- glmer(initiate.repro ~ site + (1|pop),
                  data=parent_pheno,
                  family = binomial())

summary(initiate_site)
Anova(initiate_site)

#Does initiating reproduction differ by elevation?
initiate_elev <- glm(initiate.repro ~ elev_m,
                       data=parent_pheno,
                       family=binomial())
summary(initiate_elev)
initiate_elev_null <- glm(initiate.repro ~ 1,
                          data=parent_pheno,
                          family=binomial())
lrtest(initiate_elev_null, initiate_elev)
```

#### Model: Does reproductive duration differ by site/elevation?

```{r}
#Do we want to include year as a random effect because we have 2 years at WL2
#Do we still see the same pattern removing WL2 2023 individuals?

init.repro <- parent_pheno %>%
  filter(initiate.repro == 1)

#does total reproductive duration differ by garden?
repro_duration_interaction <- lmer(total.repro.duration ~ site*elev_m + (1|pop) + (site|block),
                       data=init.repro)
summary(repro_duration_interaction)
Anova(repro_duration_interaction)

repro_duration_add <- lmer(total.repro.duration ~ site + elev_m + (1|pop) + (site|block),
                       data=init.repro)
summary(repro_duration_add)

Anova(repro_duration_add)

ICtab(repro_duration_interaction, repro_duration_add)




#does budding duration differ by garden?
bud_duration_site<- lmer(bud.duration ~ site + (1|pop),
                     data=init.repro)
summary(bud_duration_site)
Anova(bud_duration_site)

bud_duration_elev<- lmer(bud.duration ~ elev_m + (1|block),
                     data=init.repro)
summary(bud_duration_elev)
Anova(bud_duration_elev)



test <- init.repro %>%
  filter(!(site == "WL2" & year == 23))

bud_test_no_23 <- lmer(bud.duration ~ site + (1|pop),
                     data=test)
summary(bud_test_no_23)
Anova(bud_test_no_23)




#does flowering duration differ by garden?
flw_duration_site <- lmer(flower.duration ~ site + (1|pop),
                          data=init.repro)
summary(flw_duration_site)
Anova(flw_duration_site)



#does fruiting duration differ by garden?
frt_duration_site <- lmer(fruit.duration ~ site + (1|pop),
                          data=init.repro)
Anova(frt_duration_site)
```

```{r}
ggplot(init.repro, aes(x=site, y=total.repro.duration, color=elev_m)) +
  geom_point() +
  geom_smooth(method="lm", aes(group=elev_m))

ggplot(init.repro, aes(x=site, y=bud.duration, color=elev_m)) +
  geom_point() +
  geom_smooth(method="lm", aes(group=elev_m))

ggplot(init.repro, aes(x=site, y=flower.duration, color=elev_m)) +
  geom_point() +
  geom_smooth(method="lm", aes(group=elev_m))

ggplot(init.repro, aes(x=site, y=fruit.duration, color=elev_m)) +
  geom_point() +
  geom_smooth(method="lm", aes(group=elev_m))

```

#### Calculating means and sem

```{r}
sem <- function(x, na.rm=FALSE) {           #for calculating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 
repro_summary <- init.repro %>%
  group_by(pop, site, elev_m) %>%
  summarize(N_repro = sum(!is.na(total.repro.duration)),
            N_bud = sum(!is.na(bud.duration)),
            mean_total_dur = mean(total.repro.duration, na.rm=(TRUE)),
            sem_total_dur = sem(total.repro.duration, na.rm=(TRUE)),
            mean_bud_dur = mean(bud.duration, na.rm=(TRUE)),
            sem_bud_dur = sem(bud.duration, na.rm=(TRUE)),
            mean_flw_dur = mean(flower.duration, na.rm=(TRUE)),
            sem_flw_dur = sem(flower.duration, na.rm=(TRUE))
            )
```

#### Rx norms for total repro duration and bud duration - only BH and TM2 

```{r}
repro_summary %>%
  filter(N_repro != 1) %>%
  ggplot(aes(x=site, y=mean_total_dur, group=pop, color=elev_m)) +
  geom_point(size=1.5) +
  geom_line(size=1) +
  geom_errorbar(aes(ymin=mean_total_dur-sem_total_dur, ymax=mean_total_dur+sem_total_dur), width=0.1) 

repro_summary %>%
  filter(N_bud != 1) %>%
  ggplot(aes(x=site, y=mean_bud_dur, group=pop, color=elev_m)) +
  geom_point(size=1.5) +
  geom_line(size=1) +
  geom_errorbar(aes(ymin=mean_bud_dur-sem_bud_dur, ymax=mean_bud_dur+sem_bud_dur), width=0.1) 
```

#### Model: Does days to budding differ by site/elevation?

```{r}
dt_bud <- lmer(bud.date_dt ~ site*elev_m + (1|pop),
               data=init.repro)
summary(dt_bud)
Anova(dt_bud) #THIS MAY BE PICKING UP THE POPS THAT ONLY BUDDED AT 1 OF THE GARDENS

dt_flw <- lmer(flower.date_dt ~ site*elev_m + (1|pop),
               data=init.repro)
summary(dt_flw)
Anova(dt_flw)

#ONLY INCLUDING POPS THAT BUDDED AT BOTH GARDENS IN THE MODEL
both_bud <- init.repro %>%
  filter(pop %in% c("BH", "TM2", "CC", "SC"))

dt_bud_both <- lmer(bud.date_dt ~ site*elev_m + (1|pop),
               data=both_bud)
summary(dt_bud_both)
Anova(dt_bud_both) #BH is the highest of the low elevation pops and has a slighlty steeper slope than the others which is probably what is driving this site*elevation interaction, elevation is now a significant predictor because BH is at a higher elevation than the others but that is because the high elevation pops are no longer considered

```

```{r}
dt_bud_summary <- init.repro %>%
  group_by(pop, site, elev_m) %>%
  summarize(N_bud = sum(!is.na(bud.date_dt)),
            mean_dt_bud = mean(bud.date_dt, na.rm=(TRUE)),
            sem_dt_bud = sem(bud.date_dt, na.rm=(TRUE))
            )

dt_bud_summary %>%
  filter(N_bud != 1) %>%
  ggplot(aes(x=site, y=mean_dt_bud, group=pop, color=elev_m)) +
  geom_point(size=1.5) +
  geom_line(size=1) +
  geom_errorbar(aes(ymin=mean_dt_bud-sem_dt_bud, ymax=mean_dt_bud+sem_dt_bud), width=0.1) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0")
```

```{r}
dt_flw_summary <- init.repro %>%
  group_by(pop, site, elev_m) %>%
  summarize(N_flw = sum(!is.na(flower.date_dt)),
            mean_dt_flw = mean(flower.date_dt, na.rm=(TRUE)),
            sem_dt_flw = sem(flower.date_dt, na.rm=(TRUE))
            )

dt_flw_summary %>%
  filter(N_flw != 1) %>%
  ggplot(aes(x=site, y=mean_dt_flw, group=pop, color=elev_m)) +
  geom_point(size=1.5) +
  geom_line(size=1) +
  geom_errorbar(aes(ymin=mean_dt_flw-sem_dt_flw, ymax=mean_dt_flw+sem_dt_flw), width=0.1) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0")
```

```{r}
dt_fr_summary <- init.repro %>%
  group_by(pop, site, elev_m) %>%
  summarize(N_fr = sum(!is.na(fruit.date_dt)),
            mean_dt_fr = mean(fruit.date_dt, na.rm=(TRUE)),
            sem_dt_fr = sem(fruit.date_dt, na.rm=(TRUE))
            )

dt_fr_summary %>%
  filter(N_fr != 1) %>%
  ggplot(aes(x=site, y=mean_dt_fr, group=pop, color=elev_m)) +
  geom_point(size=1.5) +
  geom_line(size=1) +
  geom_errorbar(aes(ymin=mean_dt_fr-sem_dt_fr, ymax=mean_dt_fr+sem_dt_fr), width=0.1) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0")
```
