---
title: "Phenology_Models"
output: html_document
date: "2025-01-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse)
library(lubridate)
library(lme4)
library(googlesheets4)
library(lmtest)
library(car)
```

```{r}
UCD22_23 <- read_csv("../input/UCD22_23_transplant.csv")

WL2_23 <- read_csv("../input/WL2_23.csv")
WL2_24 <- read_csv("../input/WL2_24.csv")

clim_dist_UCD22_23 <- read_csv("../input/growthseason_GowersEnvtalDist_UCD.csv")
clim_dist_WL2_23 <- read_csv("../input/growthseason_GowersEnvtalDist_WL2.csv")
```

#### Select only the columns of interest and only parent pops from the WL2_24 garden

```{r}
UCD22_23 <- UCD22_23 %>%
  mutate(bed = str_extract(block, "^[A-Za-z]")) %>% #create column for bed by extracting info from block column
  select(block, bed, row, col, everything()) %>% #reorder columns
  select(block, bed, row, col, pop, mf, rep, bud.date_dfp, flower.date_dfp, fruit.date_dfp, last.FL.date_dfp, last.FR.date_dfp, death.date_dfp) %>% #select columns of interest
  filter(pop != "buffer") #filter out buffer plants

WL2_23 <- WL2_23 %>%
  select(block, bed, row, col, pop, mf, rep, bud.date_dfp, flower.date_dfp, fruit.date_dfp, last.FL.date_dfp, last.FR.date_dfp, death.date_dfp) %>%
  filter(grepl("^\\d+$", rep)) #removes buffer plants and empty spaces

WL2_24 <- WL2_24 %>%
  select(bed, row, col, pop, mf, rep, bud.date_dfp, flower.date_dfp, fruit.date_dfp, last.FL.date_dfp, last.FR.date_dfp, death.date_dfp) %>%
  left_join(WL2_23 %>% select(block, bed, row, col), by=c("bed", "row", "col")) %>% #add block info using WL2_23 block column
  select(block, bed, row, col, everything()) #reorder columns 
```

#### Combine all gardens into one tibble, add a site column

```{r}
parent_pheno <- bind_rows(
  UCD22_23 %>% mutate(site="UCD22_23"),
  WL2_23 %>% mutate(site="WL2_23"),
  WL2_24 %>% mutate(site="WL2_24")
)
```

#### Add elevation data

```{r}
clim_dist_UCD22_23 <- clim_dist_UCD22_23 %>%
  filter(TimePd == "Recent") %>%
  rename(pop=parent.pop)

parent_pheno <- parent_pheno %>%
  left_join(clim_dist_UCD22_23 %>% select(pop, elev_m), by="pop")
```

#### Calculate duration of each reproductive stage and add binary column for initiating reproduction

```{r}
#Bud duration (flower.date - bud.date)
#Flower duration (last.FL.date - flower.date)
#Fruit duration (last.FR.date - fruit.date)
#Total repro duration (last.FR.date - bud.date)
#Initiate repro(1 if there is any repro date, 0 if no repro date)
  # I would want to use bud.date to determine if a plant initiated repro, but we missed the bud date for some plants

parent_pheno <- parent_pheno %>%
  mutate(bud.duration = as.numeric(flower.date_dfp - bud.date_dfp)) %>%
  mutate(flower.duration = as.numeric(last.FL.date_dfp - flower.date_dfp)) %>%
  mutate(fruit.duration = as.numeric(last.FR.date_dfp - fruit.date_dfp)) %>%
  mutate(total.repro.duration = as.numeric(last.FR.date_dfp - bud.date_dfp)) %>%
  mutate(initiate.repro = if_else(
      !is.na(bud.date_dfp) | !is.na(flower.date_dfp) | 
      !is.na(fruit.date_dfp) | !is.na(last.FR.date_dfp) | 
      !is.na(last.FL.date_dfp), 
      1, 0
    ))
```

##### NEED TO DECIDE IF WE FILL IN FLOWERING DATE THAT WAS MISSED FOR WL2 on 7/23/2024

#### Visualizing each repro duration across sites

```{r}
parent_pheno %>%
  filter(!is.na(bud.duration)) %>%
  filter(!(pop %in% c("CP2", "SQ1", "SQ3", "WR"))) %>%
  ggplot(aes(y=bud.duration, fill=site)) +
  geom_boxplot() +
  facet_wrap(~pop,  scales="free")

parent_pheno %>%
  filter(!is.na(flower.duration)) %>%
  #filter(!(pop %in% c("CP2", "SQ1", "SQ3", "WR"))) %>%
  ggplot(aes(y=flower.duration, fill=site)) +
  geom_boxplot() +
  facet_wrap(~pop,  scales="free")

parent_pheno %>%
  filter(!is.na(fruit.duration)) %>%
  #filter(!(pop %in% c("CP2", "SQ1", "SQ3", "WR"))) %>%
  ggplot(aes(y=fruit.duration, fill=site)) +
  geom_boxplot() +
  facet_wrap(~pop,  scales="free")

parent_pheno %>%
  filter(!is.na(total.repro.duration)) %>%
  #filter(!(pop %in% c("CP2", "SQ1", "SQ3", "WR"))) %>%
  ggplot(aes(y=total.repro.duration, fill=site)) +
  geom_boxplot() +
  facet_wrap(~pop,  scales="free")
```

#### Model: Does reproductive duration differ by site/elevation?

```{r}
#does total reproductive duration differ by garden?
repro_duration <- lmer(total.repro.duration ~ site*elev_m + (1|pop) + (site|block),
                       data=parent_pheno)
summary(repro_duration)
Anova(repro_duration)

#does budding duration differ by garden?
bud_duration_site<- lmer(bud.duration ~ site + (1|pop),
                     data=parent_pheno)
summary(bud_duration_site)
Anova(bud_duration_site)

bud_duration_elev<- lmer(bud.duration ~ elev_m + (1|block),
                     data=parent_pheno)
summary(bud_duration_elev)
Anova(bud_duration_elev)

#does flowering duration differ by garden?
flw_duration_site <- lmer(flower.duration ~ site + (1|pop),
                          data=parent_pheno)
summary(flw_duration_site)
Anova(flw_duration_site)
```

#### Model: Does initiating reproduction differ by site/elevation?

```{r}
#Does initiating reproduction differ by site with pop as a random effect
initiate_site <- glmer(initiate.repro ~ site + (1|pop),
                  data=parent_pheno,
                  family = binomial())

summary(initiate_site)
Anova(initiate_site)

#Does initiating reproduction differ by elevation?
initiate_elev <- glm(initiate.repro ~ elev_m,
                       data=parent_pheno,
                       family=binomial())
summary(initiate_elev)
initiate_elev_null <- glm(initiate.repro ~ 1,
                          data=parent_pheno,
                          family=binomial())
lrtest(initiate_elev_null, initiate_elev)
```
