---
title: "Preliminary Data Exploration"
output:
  word_document: default
  html_document: default
date: "2024-10-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse)
library(lubridate)
library(lme4)
library(googlesheets4)
library(lmtest)
library(car)
```

### UCD 2022-2023 Transplants 

#### Load in mort/pheno data

```{r}
UCD_pheno_transplant <- read_csv("../input/UCD_2023_Data/UCD_transplants_pheno_mort_20231016_corrected.csv")
```

#### Renaming columns to remove spaces between words and change column name to match other gardens

```{r}
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date First Bud"] <- "bud.date"
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date First Fruit"] <- "fruit.date"
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date First Flower"] <- "flower.date"
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date Last Flower"] <- "last.FR.date"
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date Last Fruit"] <- "last.FL.date"
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Death Date"] <- "death.date"

```

#### Apply lubridate

```{r}
UCD_pheno_transplant$bud.date <- mdy(UCD_pheno_transplant$bud.date)
UCD_pheno_transplant$flower.date <- mdy(UCD_pheno_transplant$flower.date)
UCD_pheno_transplant$fruit.date <- mdy(UCD_pheno_transplant$fruit.date)
UCD_pheno_transplant$last.FL.date <- mdy(UCD_pheno_transplant$last.FL.date)
UCD_pheno_transplant$last.FR.date <- mdy(UCD_pheno_transplant$last.FR.date)
UCD_pheno_transplant$death.date <- mdy(UCD_pheno_transplant$death.date)
```

##### Identify which row has the wrong year

```{r}
# one first bud date seems to be 2022 which is not correct
bud2022 <- UCD_pheno_transplant[year(UCD_pheno_transplant$bud.date) == 2022,]
print(bud2022)

# Directly update the date of the 2022 entry to 2023
UCD_pheno_transplant[UCD_pheno_transplant$block == "L1" & 
                       UCD_pheno_transplant$row == 4 & 
                       UCD_pheno_transplant$col == "A", "bud.date"] <- 
  update(UCD_pheno_transplant[UCD_pheno_transplant$block == "L1" & 
                                UCD_pheno_transplant$row == 4 & 
                                UCD_pheno_transplant$col == "A", "bud.date"][[1]], 
         year = 2023)

# Verify the updated date
print(UCD_pheno_transplant[UCD_pheno_transplant$block == "L1" & 
                             UCD_pheno_transplant$row == 4 & 
                             UCD_pheno_transplant$col == "A", ])
```

#### Change date to days since planting

```{r}
UCD_pheno_transplant <- UCD_pheno_transplant %>%
  mutate(
    planting_date = mdy("11/27/2022") #the date plants were brought to the screenhouse
  )

UCD_pheno_transplant <- UCD_pheno_transplant %>%
  mutate(across(ends_with(".date"), ~ as.integer(as.Date(.) - planting_date), .names="{.col}_dt"))

write.csv(UCD_pheno_transplant, "../input/UCD22_23_transplant.csv", row.names = FALSE)
```

#### Visualize as bud/flower/fruit day from planting

```{r}
ggplot(UCD_pheno_transplant, aes(x=bud.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Bud")

ggplot(UCD_pheno_transplant, aes(x=flower.date_dt, fill=pop))+
  geom_histogram() +
  xlab("Days since planting, Flower")

ggplot(UCD_pheno_transplant, aes(x=fruit.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Fruit")
```

#### What happened to the plants that initiated reproduction but never finished? - looks to be a mix of death before reaching the next stage (4/10/23, 5/1/23, and 5/8/23 were main death dates) but unsure why there was death (notes do not indicate drought or other event) and just not progressing to the next stage

##### Add death date

```{r}
UCD23_May_mort <- read_csv("../input/UCD_2023_Data/Mortality_survey_20230508_corrected.csv")

UCD23_May_mort <- UCD23_May_mort %>%
  rename(death_date = `death date`)
```

##### Merge death data into phenology

```{r}
UCD23_pheno_mort <- UCD_pheno_transplant %>%
  left_join(select(UCD23_May_mort, block, row, col, death_date, Notes), by=c("block"="block", "row"="row", "col"="col"))
```

#### Calculate proportions of total planted for each reproductive stage by pop

```{r}
UCD_trans_prop <- UCD_pheno_transplant %>%
  filter(pop != "buffer") %>%
  group_by(pop) %>%
  summarise(
    total_individuals = n(),
    prop_bud = sum(!is.na(bud.date))/total_individuals,
    prop_flower = sum(!is.na(flower.date))/total_individuals,
    prop_fruit = sum(!is.na(fruit.date))/total_individuals
  ) %>%
  pivot_longer(cols=starts_with("prop"),
               names_to="stage",
               values_to="proportion")

write.csv(UCD_trans_prop, "../input/UCD22_23_transplant_prop.csv", row.names = FALSE)
```

#### Only include populations with non-zero proportion for bud date

```{r}
#UCD_trans_prop_non_zero_bud <- UCD_trans_prop %>%
  #filter(stage == "prop_bud" & proportion > 0) %>%
 # pull(pop)

#UCD_trans_prop_filter <- UCD_trans_prop %>%
  #filter(pop %in% UCD_trans_prop_non_zero_bud)
```

#### Visualize proportions

```{r}
ggplot(UCD_trans_prop, aes(x=stage, y=proportion, fill=stage)) + 
  geom_bar(stat="identity", show.legend=FALSE) +
  geom_text(aes(label=paste0("Total: ", total_individuals)),
            position = position_stack(vjust = 1.1), size = 2, color = "black") +
  facet_wrap(~pop, ncol = 4) +
  labs(x="Reproductive Stage", y="Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### UCD 2022-2023 Germinants - DON'T NEED TO RUN

```{r}
UCD_pheno_germ <- read_csv("../input/UCD_2023_Data/Phenology_survey_germination_20230515_corrected.csv")
```

#### Renaming columns to remove spaces between words and apply lubridate

```{r}
#Update column names
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date First Bud"] <- "bud.date"
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date First Fruit"] <- "fruit.date"
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date First Flower"] <- "flower.date"
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date Last Flower"] <- "last.FL.date"
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date Last Fruit"] <- "last.FR.date"

#apply lubridate
UCD_pheno_germ$bud.date <- mdy(UCD_pheno_germ$bud.date)
UCD_pheno_germ$flower.date <- mdy(UCD_pheno_germ$flower.date)
UCD_pheno_germ$fruit.date <- mdy(UCD_pheno_germ$fruit.date)
UCD_pheno_germ$last.FL.date <- mdy(UCD_pheno_germ$last.FL.date) #no last flower dates
UCD_pheno_germ$last.FR.date <- mdy(UCD_pheno_germ$last.FR.date)
```

#### Change date to days since planting

```{r}
UCD_pheno_germ <- UCD_pheno_germ %>%
  mutate(
    planting_date = mdy("11/27/2022")
  )

UCD_pheno_germ <- UCD_pheno_germ %>%
  mutate(across(ends_with(".date"), ~ as.integer(. - planting_date), .names="{.col}_dfp"))

write.csv(UCD_pheno_germ, "../input/UCD22_23_germ.csv", row.names=FALSE)
```

#### Visualize as bud/flower/fruit day from planting

```{r}
ggplot(UCD_pheno_germ, aes(x=bud.date_dfp, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Bud")

ggplot(UCD_pheno_germ, aes(x=flower.date_dfp, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Flower")

ggplot(UCD_pheno_germ, aes(x=fruit.date_dfp, fill=pop)) +
  geom_histogram() +
   xlab("Days since planting, Fruit")
```

#### Calculate proportion of total planted for each reproductive stage by pop

```{r}
UCD_germ_prop <- UCD_pheno_germ %>%
  group_by(pop) %>%
  summarise(
    total_individuals = n(),
    prop_bud = sum(!is.na(bud.date))/total_individuals,
    prop_flower = sum(!is.na(flower.date))/total_individuals,
    prop_fruit = sum(!is.na(fruit.date))/total_individuals
  ) %>%
  pivot_longer(cols=starts_with("prop"),
               names_to="stage",
               values_to="proportion")

write.csv(UCD_germ_prop, "../input/UCD22_23_germ_prop.csv", row.names = FALSE)
```

#### Visualize proportions

```{r}
ggplot(UCD_germ_prop, aes(x=stage, y=proportion, fill=stage)) + 
  geom_bar(stat="identity", show.legend=FALSE) +
  geom_text(aes(label=paste0("Total: ", total_individuals)),
            position = position_stack(vjust = 1.1), size = 2, color = "black") +
  facet_wrap(~pop, ncol = 4) +
  labs(x="Reproductive Stage", y="Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Visualize proportion for just TM2

```{r}
UCD_germ_prop %>%
  filter(pop == "TM2") %>%
  ggplot(aes(x=stage, y=proportion, fill=stage)) + 
  geom_bar(stat="identity", show.legend=FALSE) +
  geom_text(aes(label=paste0("Total: ", total_individuals)),
            position = position_stack(vjust = 1.1), size = 2, color = "black") +
  facet_wrap(~pop, ncol = 4) +
  labs(x="Reproductive Stage", y="Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### UCD 2023-2024

#### Load in data

```{r}
UCD24 <- read_csv("../input/UCD_2023_2024_Data/UCD_mort_pheno_20240908_corrected.csv")

#need to get pop info from uniqueIDs
url <- "https://docs.google.com/spreadsheets/d/1BM5wVLOffxUdHC1yCBJbv3le5hf7Q-WspYVM_nYBiXQ/edit?gid=123887845#gid=123887845"
parent_pop_info <- read_sheet(url, sheet="parent pop labels and germination")

```

#### Add pop column for parent pops using unique.ID

```{r}
UCD24_pop <- UCD24 %>%
  left_join(
    parent_pop_info %>%
      select(unique.ID, pop.id) %>%
      mutate(unique.ID = as.character(unique.ID)),  # Convert unique.ID to character
    by = "unique.ID"
  )

colnames(UCD24_pop)[colnames(UCD24_pop) == "pop.id"] <- "pop"
```

#### Apply lubridate

```{r}
UCD24_pop$bud.date <- mdy(UCD24_pop$bud.date)
UCD24_pop$flower.date <- mdy(UCD24_pop$flower.date)
UCD24_pop$fruit.date <- mdy(UCD24_pop$fruit.date)
UCD24_pop$last.FL.date <- mdy(UCD24_pop$last.FL.date)
UCD24_pop$last.FR.date <- mdy(UCD24_pop$last.FR.date)
UCD24_pop$death.date <- mdy(UCD24_pop$death.date)
UCD24_pop$missing.date <- mdy(UCD24_pop$missing.date)

```

#### Fix incorrect dates

```{r}
bud2021 <- UCD24_pop[year(UCD24_pop$bud.date) == 2021,]
print(bud2021)

# Directly update the date of the 2021 entry
UCD24_pop[UCD24_pop$bed == "A" & 
                       UCD24_pop$row == 15 & 
                       UCD24_pop$col == "A", "bud.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "A" & 
                                UCD24_pop$row == 15 & 
                                UCD24_pop$col == "A", "bud.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "A" & 
                             UCD24_pop$row == 15 & 
                             UCD24_pop$col == "A", ])

bud2014 <- UCD24_pop[year(UCD24_pop$bud.date) == 2014,]
print(bud2014)

# Directly update the date of the 2014 entry
UCD24_pop[UCD24_pop$bed == "C" & 
                       UCD24_pop$row == 35 & 
                       UCD24_pop$col == "D", "bud.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "C" & 
                                UCD24_pop$row == 35 & 
                                UCD24_pop$col == "D", "bud.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "C" & 
                             UCD24_pop$row == 35 & 
                             UCD24_pop$col == "D", ])

flw2023 <- UCD24_pop[year(UCD24_pop$flower.date) == 2023,]
print(flw2023)

# Directly update the date of the 2023 entry
UCD24_pop[UCD24_pop$bed == "A" & 
                       UCD24_pop$row == 17 & 
                       UCD24_pop$col == "C", "flower.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "A" & 
                                UCD24_pop$row == 17 & 
                                UCD24_pop$col == "C", "flower.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "A" & 
                             UCD24_pop$row == 17 & 
                             UCD24_pop$col == "C", ])

# Directly update the date of the 2023 entry
UCD24_pop[UCD24_pop$bed == "A" & 
                       UCD24_pop$row == 17 & 
                       UCD24_pop$col == "C", "fruit.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "A" & 
                                UCD24_pop$row == 17 & 
                                UCD24_pop$col == "C", "fruit.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "A" & 
                             UCD24_pop$row == 17 & 
                             UCD24_pop$col == "C", ])

# Directly update the date of the 2014 entry
UCD24_pop[UCD24_pop$bed == "B" & 
                       UCD24_pop$row == 14 & 
                       UCD24_pop$col == "A", "fruit.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "B" & 
                                UCD24_pop$row == 14 & 
                                UCD24_pop$col == "A", "fruit.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "B" & 
                             UCD24_pop$row == 14 & 
                             UCD24_pop$col == "A", ])

# Directly update the date of the 2014 entry
UCD24_pop[UCD24_pop$bed == "D" & 
                       UCD24_pop$row == 44 & 
                       UCD24_pop$col == "D", "last.FL.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "D" & 
                                UCD24_pop$row == 44 & 
                                UCD24_pop$col == "D", "last.FL.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "D" & 
                             UCD24_pop$row == 44 & 
                             UCD24_pop$col == "D", ])

#CHECK THIS ONE, LAST FR DATE BEFORE LAST FLOWER DATE!!!! - corrected
#unique.ID 246
#Checked pdf and confirmed last FR date should be 05/31/2024

# Directly update the date of the 2014 entry
UCD24_pop[UCD24_pop$bed == "F" & 
                       UCD24_pop$row == 33 & 
                       UCD24_pop$col == "C", "last.FR.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "F" & 
                                UCD24_pop$row == 33 & 
                                UCD24_pop$col == "C", "last.FR.date"][[1]], 
         year = 2024, day = 31)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "F" & 
                             UCD24_pop$row == 33 & 
                             UCD24_pop$col == "C", ])

# Directly update the date of the 2014 entry
UCD24_pop[UCD24_pop$bed == "C" & 
                       UCD24_pop$row == 31 & 
                       UCD24_pop$col == "C", "last.FR.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "C" & 
                                UCD24_pop$row == 31 & 
                                UCD24_pop$col == "C", "last.FR.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "C" & 
                             UCD24_pop$row == 31 & 
                             UCD24_pop$col == "C", ])

#correct last fruit date
UCD24_pop[UCD24_pop$bed == "F" &
            UCD24_pop$row == 34 &
            UCD24_pop$col == "D", "last.FR.date"] <-
  update(UCD24_pop[UCD24_pop$bed == "F" &
            UCD24_pop$row == 34 &
            UCD24_pop$col == "D", "last.FR.date"][[1]],
  month=05)
```

#### Change date to days since planting

```{r}
UCD23_24_planting_date <- mdy("12/04/2023")

UCD24_pop <- UCD24_pop %>%
  mutate(across(ends_with(".date") & !matches("census.date"),
                ~ as.integer(ymd(.) - UCD23_24_planting_date),
                .names="{.col}_dt"))

write.csv(UCD24_pop, "../input/UCD23_24.csv", row.names=FALSE)
```

#### Visualize with day since planting

```{r}
UCD24_pop %>%
  filter(!is.na(pop)) %>%
  ggplot(aes(x=bud.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Bud")

UCD24_pop %>%
  filter(!is.na(pop)) %>%
  ggplot(aes(x=flower.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Flower")

UCD24_pop %>%
  filter(!is.na(pop)) %>%
  ggplot(aes(x=fruit.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Fruit")

UCD24_pop %>%
  filter(!is.na(pop)) %>%
  ggplot(aes(x=last.FL.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Last Flower")

UCD24_pop %>%
  filter(!is.na(pop)) %>%
  ggplot(aes(x=last.FR.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Last Fruit")

UCD24_pop %>%
  filter(!is.na(pop)) %>%
  ggplot(aes(x=death.date_dt, fill=pop)) +
  geom_histogram()
```

#### Calculate proportions

##### Check for cases when flower date is NA but there is a fruit date- found 1 and updated flowering date to be fruit date based on notes on datasheet

```{r}
# Find individuals with a fruit date but no flower date 
UCD24_fruit_no_flower <- UCD24_pop[!is.na(UCD24_pop$fruit.date) & is.na(UCD24_pop$flower.date), ]  

# Display the result for this condition 
print(UCD24_fruit_no_flower)   

#Add flowering date (4/26/2024) that is the same as the fruit date 
UCD24_pop$flower.date[UCD24_pop$pop == "TM2" & !is.na(UCD24_pop$fruit.date) & is.na(UCD24_pop$flower.date)] <- "2024-04-26"

```

##### Check for last fruit date but no fruit date

```{r}
UCD24_lastfr_no_fruit <- UCD24_pop[!is.na(UCD24_pop$last.FR.date) & is.na(UCD24_pop$fruit.date), ] 
print(UCD24_lastfr_no_fruit) 

#only one is parent pop - will use last fruit date as fruit date since no notes on datascan
UCD24_pop$fruit.date[UCD24_pop$pop == "WL2" & !is.na(UCD24_pop$last.FR.date) & is.na(UCD24_pop$fruit.date)] <- "2024-06-07"
```

```{r}
UCD_24_proportion <- UCD24_pop %>%
  group_by(pop) %>%
  summarise(
    total_individuals = n(),
    prop_bud = sum(!is.na(bud.date))/total_individuals,
    prop_flower = sum(!is.na(flower.date))/total_individuals,
    prop_fruit = sum(!is.na(fruit.date))/total_individuals
  ) %>%
  pivot_longer(cols=starts_with("prop"),
               names_to="stage",
               values_to="proportion") %>%
  filter(!is.na(pop)) %>%
  ungroup()

write.csv(UCD_24_proportion, "../input/UCD23_24_prop.csv", row.names = FALSE)
```

#### Visualize proportions

```{r}
UCD_24_proportion %>%
  filter(!is.na(pop)) %>%
  ggplot(aes(x=stage, y=proportion, fill=stage)) + 
  geom_bar(stat="identity", show.legend=FALSE) +
  geom_text(aes(label=paste0("Total: ", total_individuals)),
            position = position_stack(vjust = 1.1), size = 2, color = "black") +
  facet_wrap(~pop, ncol = 4, scales= "free") +
  labs(x="Reproductive Stage", y="Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Searching for any flowering props that are lower than fruiting props

```{r}
# Filter data for 'prop_flower' and 'prop_fruit' stages 
flower_data <- UCD_24_proportion[UCD_24_proportion$stage == "prop_flower", ] 
fruit_data <- UCD_24_proportion[UCD_24_proportion$stage == "prop_fruit", ]  

# Merge the data on 'pop' 
merged_data <- merge(flower_data, fruit_data, by = "pop", suffixes = c("_flower", "_fruit"))  

# Check where 'prop_flower' is smaller than 'prop_fruit' 

result <- merged_data[merged_data$proportion_flower < merged_data$proportion_fruit, ]  

# Display the result 
print(result[, c("pop", "proportion_flower", "proportion_fruit")]) 
```

### WL2 2023

#### Load in data

```{r}
WL2_pheno <- read_csv("../input/WL2_2023_Data/WL2_mort_pheno_20231020_corrected.csv")
```

#### Change column names to match other data sets

```{r}
colnames(WL2_pheno)[colnames(WL2_pheno) == "bed.row"] <- "row"
colnames(WL2_pheno)[colnames(WL2_pheno) == "bed.col"] <- "col"
colnames(WL2_pheno)[colnames(WL2_pheno) == "last.flower.date"] <- "last.FL.date"
colnames(WL2_pheno)[colnames(WL2_pheno) == "last.fruit.date"] <- "last.FR.date"
```

#### Apply lubridate

```{r}
WL2_pheno$bud.date <- mdy(WL2_pheno$bud.date)
WL2_pheno$flower.date <- mdy(WL2_pheno$flower.date)
WL2_pheno$fruit.date <- mdy(WL2_pheno$fruit.date)
WL2_pheno$last.FL.date <- mdy(WL2_pheno$last.FL.date)
WL2_pheno$last.FR.date <- mdy(WL2_pheno$last.FR.date)
WL2_pheno$death.date <- mdy(WL2_pheno$death.date)
```

#### One individual has the wrong fruit date year, correct to say 2023

```{r}
#one of the fruit dates says 2013
fruit2014 <- WL2_pheno[year(WL2_pheno$fruit.date) == 2013,]
print(fruit2014)

# Directly update the date of the 2013 entry
WL2_pheno[WL2_pheno$block == "A" & 
                       WL2_pheno$bed == "A" & 
                       WL2_pheno$row == 23 & WL2_pheno$col == "D", "fruit.date"] <- 
  update(WL2_pheno[WL2_pheno$block == "A" & 
                       WL2_pheno$bed == "A" & 
                       WL2_pheno$row == 23 & WL2_pheno$col == "D", "fruit.date"][[1]], year = 2023)

# Verify the updated date
print(WL2_pheno[WL2_pheno$block == "A" & 
                       WL2_pheno$bed == "A" & 
                       WL2_pheno$row == 23 & WL2_pheno$col == "D", ])


#UPDATE POP name for one YO4 and one YO8
WL2_pheno <- WL2_pheno %>%
  mutate(pop = case_when(
    pop == "Y04" ~ "YO4",
    pop == "Y08" ~ "YO8",
    TRUE ~ pop
  ))
```

```{r}
#THIS PLANTING IS HARDER TO ASSIGN A PLANTING DAY SINCE IT IS DONE BY BLOCK
#I will call the planting date the day when all the plants were in field conditions the previous year
WL2_23_planting_date <- mdy("07/19/2023")

WL2_pheno <- WL2_pheno %>%
  mutate(across(ends_with(".date"),
                ~ as.integer(ymd(.) - WL2_23_planting_date),
                .names="{.col}_dt"))

WL2_pheno <- WL2_pheno %>%
  mutate(mf= as.numeric(mf),
         rep= as.numeric(rep))

write.csv(WL2_pheno, "../input/WL2_23.csv", row.names = FALSE)
```

#### Visualize stage from planting date

```{r}
ggplot(WL2_pheno, aes(x=bud.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Bud")

ggplot(WL2_pheno, aes(x=flower.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Flower")

ggplot(WL2_pheno, aes(x=fruit.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Fruit")

ggplot(WL2_pheno, aes(x=death.date_dt, fill=pop)) +
  geom_histogram()
```

#### Calculate proportion

```{r}
WL2_pheno_proportion <- WL2_pheno %>%
  group_by(pop) %>%
  summarise(
    total_individuals = n(),
    prop_bud = sum(!is.na(bud.date))/total_individuals,
    prop_flower = sum(!is.na(flower.date))/total_individuals,
    prop_fruit = sum(!is.na(fruit.date))/total_individuals
  ) %>%
  pivot_longer(cols=starts_with("prop"),
               names_to="stage",
               values_to="proportion")

write.csv(WL2_pheno_proportion, "../input/WL2_23_prop.csv", row.names = FALSE)
```

#### Visualize proportions

```{r}
ggplot(WL2_pheno_proportion, aes(x=stage, y=proportion, fill=stage)) + 
  geom_bar(stat="identity", show.legend=FALSE) +
  geom_text(aes(label=paste0("Total: ", total_individuals)),
            position = position_stack(vjust = 1.1), size = 2, color = "black") +
  facet_wrap(~pop, ncol = 4) +
  labs(x="Phenological Stage", y="Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
WL2_pheno_proportion %>%
  filter(pop == "TM2" | pop == "FR") %>%
  ggplot(aes(x=stage, y=proportion, fill=stage)) + 
  geom_bar(stat="identity", show.legend=FALSE) +
  geom_text(aes(label=paste0("Total: ", total_individuals)),
            position = position_stack(vjust = 1.1), size = 2, color = "black") +
  facet_wrap(~pop, ncol = 4) +
  labs(x="Reproductive Stage", y="Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
```

### WL2 2024

#### Load in data

```{r}
WL2_24 <- read_csv("../input/WL2_2024_Data/WL2_mort_pheno_20241023_corrected.csv")

```

#### Apply lubridate

```{r}
WL2_24$bud.date <- mdy(WL2_24$bud.date)
WL2_24$flower.date <- mdy(WL2_24$flower.date)
WL2_24$fruit.date <- mdy(WL2_24$fruit.date)
WL2_24$last.FL.date <- mdy(WL2_24$last.FL.date)
WL2_24$last.FR.date <- mdy(WL2_24$last.FR.date)
WL2_24$death.date <- mdy(WL2_24$death.date)
WL2_24$missing.date <- mdy(WL2_24$missing.date)
```

```{r}
#what do I do about this? set last fl the same as last fr? YES notes confirm this
#Update missing last flower dates here
WL2_24$last.FL.date[WL2_24$last.FL.date[WL2_24$bed == "F" & WL2_24$row == 23 & WL2_24$col == "C"]] <-  as.Date("2024-08-27")
```

#### Adding pop info for only 2nd year parent pop plants

```{r}
#keep only 2nd year parent plant rows
WL2_24_pop <- WL2_24 %>%
  filter(grepl("^[A-Za-z]", unique.ID)) %>% #parent plants start with a letter
  filter(unique.ID != "buffer") #buffers also start with a letter so we remove them

#separate the unique.ID column into seperate pop, mf, and rep columns
WL2_24_pop <- WL2_24_pop %>%
  mutate(
    pop = case_when(
      grepl("^\\w+_\\d+_\\d+$", unique.ID) ~ sub("^(\\w+)_\\d+_\\d+$", "\\1", unique.ID),
      TRUE ~ unique.ID
    ),
    mf = case_when(
      grepl("^\\w+_\\d+_\\d+$", unique.ID) ~ sub("^\\w+_(\\d+)_\\d+$", "\\1", unique.ID),
      TRUE ~ unique.ID
    ),
    rep = case_when(
      grepl("^\\w+_\\d+_\\d+$", unique.ID) ~ sub("^\\w+_\\d+_(\\d+)$", "\\1", unique.ID),
      TRUE ~ unique.ID
    )
  )

```

#### Update missing flower dates

```{r}
WL2_24_pop$flower.date[WL2_24_pop$bed == "A" & WL2_24_pop$row == 18 & WL2_24_pop$col == "A"] <- as.Date("2024-07-23")

WL2_24_pop$flower.date[WL2_24_pop$bed == "E" & WL2_24_pop$row == 49 & WL2_24_pop$col == "B"] <- as.Date("2024-07-23")

WL2_24_pop$flower.date[WL2_24_pop$bed == "J" & WL2_24_pop$row == 4 & WL2_24_pop$col == "A"] <- as.Date("2024-07-09")

WL2_24_pop_no_flw <- WL2_24_pop %>%
  filter(!is.na(fruit.date) & is.na(flower.date)) #only individual with a fruit date missing a flower date is a plant mistakenly coded as a buffer
```

#### Convert date to days since planting

```{r}
#WL2_24_pop <- WL2_24_pop %>%
  #mutate(
   # planting_date = if_else(is.na(block), WL2_23_planting_date, NA_Date_)
 # ) %>%
 # mutate(across(
  #  ends_with(".date"),  
  #  ~ if_else(is.na(block), as.integer(ymd(.) - WL2_23_planting_date), NA_integer_),  
  #  .names = "{.col}_dfp"  
 # ))

#ALSO CALCULATE DAYS SINCE SNOW MELT TO GET WITHIN GROWING SEASON TIMING
WL2_24_snow_melt <- mdy("05/19/2024")

WL2_24_pop <- WL2_24_pop %>%
  mutate(across(ends_with(".date"),
                ~ as.integer(ymd(.) - WL2_24_snow_melt),
                .names="{.col}_dt"))

write.csv(WL2_24_pop, "../input/WL2_24.csv", row.names=FALSE)
```

#### Visualize days since planting

```{r}

ggplot(WL2_24_pop, aes(x=bud.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Bud")

ggplot(WL2_24_pop, aes(x=flower.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Flower")

ggplot(WL2_24_pop, aes(x=fruit.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Fruit")

ggplot(WL2_24_pop, aes(x=last.FL.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Last Flower")

ggplot(WL2_24_pop, aes(x=last.FR.date_dt, fill=pop)) +
  geom_histogram() +
  xlab("Days since planting, Last Fruit")
```

#### Calculate proportion

##### Finding plants with a flower date and NA bud date/fruit date and NA flower date

```{r}
#finding plants that have a fruit date but no flower date
WL2_fruit_no_flower <- WL2_24_pop[!is.na(WL2_24_pop$fruit.date) & is.na(WL2_24_pop$flower.date), ]

# View the rows
print(WL2_fruit_no_flower)

#updating the flower date to be the same as the fruit date after confirming notes in datascans
WL2_24_pop$flower.date[is.na(WL2_24_pop$flower.date) & !is.na(WL2_24_pop$fruit.date)] <- WL2_24_pop$fruit.date[is.na(WL2_24_pop$flower.date) & !is.na(WL2_24_pop$fruit.date)]

# Verify the update
print(WL2_24_pop[is.na(WL2_24_pop$flower.date) & !is.na(WL2_24_pop$fruit.date), ])


#finding plants that have a flower date but no bud date
WL2_flower_no_bud <- WL2_24_pop[!is.na(WL2_24_pop$flower.date) & is.na(WL2_24_pop$bud.date), ]

# View the rows
print(WL2_flower_no_bud)

#updating budding date to be the same as flowering date for all but WL2_8_4 in bed B, row 26, col A after reviewing the notes from the datascan
# Identify rows with a flower date but an NA for bud date
rows_to_update <- WL2_24_pop[!is.na(WL2_24_pop$flower.date) & is.na(WL2_24_pop$bud.date), ]

# Update the bud_date to be the same as flower_date for these rows, except the specific row (bed == "F", row == 26, col == "A")
WL2_24_pop$bud.date[!is.na(WL2_24_pop$flower.date) & is.na(WL2_24_pop$bud.date) & 
                             !(WL2_24_pop$bed == "F" & WL2_24_pop$row == 26 & WL2_24_pop$col == "A")] <- 
                             WL2_24_pop$flower.date[!is.na(WL2_24_pop$flower.date) & is.na(WL2_24_pop$bud.date) & 
                             !(WL2_24_pop$bed == "F" & WL2_24_pop$row == 26 & WL2_24_pop$col == "A")]

# For the specific row where bed == "F", row == 26, and col == "A", set bud_date to be 7 days before flower_date
WL2_24_pop$bud.date[WL2_24_pop$bed == "F" & WL2_24_pop$row == 26 & WL2_24_pop$col == "A"] <- 
  as.Date(WL2_24_pop$flower.date[WL2_24_pop$bed == "F" & WL2_24_pop$row == 26 & WL2_24_pop$col == "A"]) - 7

# Verify the update
print(WL2_24_pop[!is.na(WL2_24_pop$flower.date) & is.na(WL2_24_pop$bud.date), ])
```

##### NEED TO ALSO MAKE SURE TO CHECK IF THERE IS A LAST FRUIT WITHOUT Fruit or last flower without flower or last fr and no last fl

```{r}
WL224_lastfr_no_fruit <- WL2_24_pop[!is.na(WL2_24_pop$last.FR.date) & is.na(WL2_24_pop$fruit.date), ] 
print(WL224_lastfr_no_fruit) 

WL224_lastfl_no_flw <- WL2_24_pop[!is.na(WL2_24_pop$last.FL.date) & is.na(WL2_24_pop$flower.date), ] 
print(WL224_lastfl_no_flw) 

WL224_lastfl_no_lastflw <- WL2_24_pop[!is.na(WL2_24_pop$last.FR.date) & is.na(WL2_24_pop$last.FL.date), ] 
print(WL224_lastfl_no_lastflw) 



```

```{r}
WL2_24_proportion <- WL2_24_pop %>%
  group_by(pop) %>%
  summarise(
    total_individuals = n(),
    prop_bud = sum(!is.na(bud.date))/total_individuals,
    prop_flower = sum(!is.na(flower.date))/total_individuals,
    prop_fruit = sum(!is.na(fruit.date))/total_individuals
  ) %>%
  pivot_longer(cols=starts_with("prop"),
               names_to="stage",
               values_to="proportion")
```

```{r}
#only inlcude parent pops?
WL2_24_prop_parent <- WL2_24_proportion %>%
  filter(grepl("^[A-Z]+[0-9]*$", pop) & pop != "buffer")

write.csv(WL2_24_prop_parent, "../input/WL2_24_prop.csv", row.names = FALSE)
```

```{r}
#only include populations with non-zero proportion for bud date
#WL2_24_prop_non_zero_bud <- WL2_24_proportion %>%
 # filter(stage == "prop_bud" & proportion > 0) %>%
 # pull(pop)

#WL2_24_prop_filter <- WL2_24_proportion %>%
  #filter(pop %in% WL2_24_prop_non_zero_bud)
```

```{r}
WL2_24_prop_parent %>%
  filter(total_individuals > 1) %>%
  ggplot(aes(x=stage, y=proportion, fill=stage)) + 
  geom_bar(stat="identity", show.legend=FALSE) +
  geom_text(aes(label=paste0("Total: ", total_individuals)),
            position = position_stack(vjust = 1.1), size = 2, color = "black") +
  facet_wrap(~pop, ncol = 4) +
  labs(x="Reproductive Stage", y="Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Searching for stage proportions that are greater than the previous stage proportion

```{r}
# Filter data for 'prop_flower' and 'prop_fruit' stages 
bud_data2 <- WL2_24_prop_parent[WL2_24_prop_parent$stage == "prop_bud", ]
flower_data2 <- WL2_24_prop_parent[WL2_24_prop_parent$stage == "prop_flower", ] 
fruit_data2 <- WL2_24_prop_parent[WL2_24_prop_parent$stage == "prop_fruit", ]  

# Merge the data on 'pop' 
merged_data2 <- merge(flower_data2, fruit_data2, by = "pop", suffixes = c("_flower", "_fruit"))  
merged_data3 <- merge(merged_data2, bud_data2, by="pop")

# Check where 'prop_flower' is smaller than 'prop_fruit' 

result2 <- merged_data2[merged_data2$proportion_flower < merged_data2$proportion_fruit, ]  
result3 <- merged_data3[merged_data3$proportion < merged_data3$proportion_flower,]

print(result2[, c("pop", "proportion_flower", "proportion_fruit")]) 
print(result3[, c("pop", "proportion", "proportion_flower")])
```

Likelihood of flowering given budding

Reaction Norm: days to flowering at high and low elevation

Are there differences in flowering timing plasticity between populations

What are the fates of WL2 plants that survived to year 2 but did not reproduce? stacked bar plot with year 2 fate
