---
title: "Preliminary Data Exploration"
output:
  word_document: default
  html_document: default
date: "2024-10-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse)
library(lubridate)
library(lme4)
library(googlesheets4)
```

### UCD 2022-2023 Transplants

```{r}
UCD_pheno_transplant <- read_csv("../input/UCD_2023_Data/Phenology_survey_transplants_20230508_corrected.csv")
```

#### Renaming columns to remove spaces between words

```{r}
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date First Bud"] <- "Date_First_Bud"
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date First Fruit"] <- "Date_First_Fruit"
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date First Flower"] <- "Date_First_Flower"
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date Last Flower"] <- "Date_Last_FLower"
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date Last Fruit"] <- "Date_Last_Fruit"

```

#### Apply lubridate

```{r}
UCD_pheno_transplant$Date_First_Bud <- mdy(UCD_pheno_transplant$Date_First_Bud)
UCD_pheno_transplant$Date_First_Flower <- mdy(UCD_pheno_transplant$Date_First_Flower)
UCD_pheno_transplant$Date_First_Fruit <- mdy(UCD_pheno_transplant$Date_First_Fruit)
#UCD_pheno_transplant$Date_Last_Flower <- mdy(UCD_pheno_transplant$Date_Last_Flower) #no last flower dates
#UCD_pheno_transplant$Date_Last_Fruit <- mdy(UCD_pheno_transplant$Date_Last_Fruit) #no last fruit dates
```

#### Filter to only include individuals that budded

```{r}
UCD_pheno_trans_filter <- UCD_pheno_transplant[!is.na(UCD_pheno_transplant$Date_First_Bud) | !is.na(UCD_pheno_transplant$Date_First_Flower), ]
```

#### Visualize bud date - the year for one individual is wrong

```{r}
ggplot(UCD_pheno_trans_filter, aes(x=Date_First_Bud)) +
  geom_histogram()
```

##### Identify which row has the wrong year

```{r}
# one first bud date seems to be 2022 which is not correct
bud2022 <- UCD_pheno_trans_filter[year(UCD_pheno_trans_filter$Date_First_Bud) == 2022,]
print(bud2022)
```

##### Update the year to 2023

```{r}
# Directly update the date of the 2022 entry
UCD_pheno_trans_filter[UCD_pheno_trans_filter$block == "L1" & 
                       UCD_pheno_trans_filter$row == 4 & 
                       UCD_pheno_trans_filter$col == "A", "Date_First_Bud"] <- 
  update(UCD_pheno_trans_filter[UCD_pheno_trans_filter$block == "L1" & 
                                UCD_pheno_trans_filter$row == 4 & 
                                UCD_pheno_trans_filter$col == "A", "Date_First_Bud"][[1]], 
         year = 2023)

# Verify the updated date
print(UCD_pheno_trans_filter[UCD_pheno_trans_filter$block == "L1" & 
                             UCD_pheno_trans_filter$row == 4 & 
                             UCD_pheno_trans_filter$col == "A", ])

```

#### Visualize bud, flower, and fruit dates

```{r}
ggplot(UCD_pheno_trans_filter, aes(x=Date_First_Bud, fill=pop)) +
  geom_histogram()

ggplot(UCD_pheno_trans_filter, aes(x=Date_First_Flower, fill=pop))+
  geom_histogram()

ggplot(UCD_pheno_trans_filter, aes(x=Date_First_Fruit, fill=pop)) +
  geom_histogram()
```

```{r}
#m1 <- glmer(Date_First_Bud ~ 1 + (1|pop),
           # family=gaussian,
           # data=UCD_pheno_transplant)
```

### UCD 2022-2023 Germinants

```{r}
UCD_pheno_germ <- read_csv("../input/UCD_2023_Data/Phenology_survey_germination_20230515_corrected.csv")
```

#### Renaming columns to remove spaces between words and apply lubridate

```{r}
#Update column names
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date First Bud"] <- "Date_First_Bud"
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date First Fruit"] <- "Date_First_Fruit"
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date First Flower"] <- "Date_First_Flower"
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date Last Flower"] <- "Date_Last_FLower"
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date Last Fruit"] <- "Date_Last_Fruit"

#apply lubridate
UCD_pheno_germ$Date_First_Bud <- mdy(UCD_pheno_germ$Date_First_Bud)
UCD_pheno_germ$Date_First_Flower <- mdy(UCD_pheno_germ$Date_First_Flower)
UCD_pheno_germ$Date_First_Fruit <- mdy(UCD_pheno_germ$Date_First_Fruit)
#UCD_pheno_germ$Date_Last_Flower <- mdy(UCD_pheno_germ$Date_Last_Flower) #no last flower dates
UCD_pheno_germ$Date_Last_Fruit <- mdy(UCD_pheno_germ$Date_Last_Fruit)
```

#### Visualize bud, flower, and fruit dates

```{r}
ggplot(UCD_pheno_germ, aes(x=Date_First_Bud, fill=pop)) +
  geom_histogram()

ggplot(UCD_pheno_germ, aes(x=Date_First_Flower, fill=pop)) +
  geom_histogram()

ggplot(UCD_pheno_germ, aes(x=Date_First_Fruit, fill=pop)) +
  geom_histogram()
```

### UCD 2023-2024

```{r}
UCD24 <- read_csv("../input/UCD_2023_2024_Data/UCD_mort_pheno_20240908_corrected.csv")

#need to get pop info from uniqueIDs
url <- "https://docs.google.com/spreadsheets/d/1BM5wVLOffxUdHC1yCBJbv3le5hf7Q-WspYVM_nYBiXQ/edit?gid=123887845#gid=123887845"
parent_pop_info <- read_sheet(url, sheet="parent pop labels and germination")

```

```{r}
#add pop column for parent pops using unique.ID
UCD24_pop <- UCD24 %>%
  left_join(
    parent_pop_info %>%
      select(unique.ID, pop.id) %>%
      mutate(unique.ID = as.character(unique.ID)),  # Convert unique.ID to character
    by = "unique.ID"
  )
```

```{r}
#apply lubridate
UCD24_pop$bud.date <- mdy(UCD24_pop$bud.date)
UCD24_pop$flower.date <- mdy(UCD24_pop$flower.date)
UCD24_pop$fruit.date <- mdy(UCD24_pop$fruit.date)
UCD24_pop$last.FL.date <- mdy(UCD24_pop$last.FL.date)
UCD24_pop$last.FR.date <- mdy(UCD24_pop$last.FR.date)
UCD24_pop$death.date <- mdy(UCD24_pop$death.date)
UCD24_pop$missing.date <- mdy(UCD24_pop$missing.date)

```

```{r}
bud2021 <- UCD24_pop[year(UCD24_pop$bud.date) == 2021,]
print(bud2021)

# Directly update the date of the 2021 entry
UCD24_pop[UCD24_pop$bed == "A" & 
                       UCD24_pop$row == 15 & 
                       UCD24_pop$col == "A", "bud.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "A" & 
                                UCD24_pop$row == 15 & 
                                UCD24_pop$col == "A", "bud.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "A" & 
                             UCD24_pop$row == 15 & 
                             UCD24_pop$col == "A", ])
```

```{r}
bud2014 <- UCD24_pop[year(UCD24_pop$bud.date) == 2014,]
print(bud2014)

# Directly update the date of the 2014 entry
UCD24_pop[UCD24_pop$bed == "C" & 
                       UCD24_pop$row == 35 & 
                       UCD24_pop$col == "D", "bud.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "C" & 
                                UCD24_pop$row == 35 & 
                                UCD24_pop$col == "D", "bud.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "C" & 
                             UCD24_pop$row == 35 & 
                             UCD24_pop$col == "D", ])
```

```{r}
flw2023 <- UCD24_pop[year(UCD24_pop$flower.date) == 2023,]
print(flw2023)

# Directly update the date of the 2023 entry
UCD24_pop[UCD24_pop$bed == "A" & 
                       UCD24_pop$row == 17 & 
                       UCD24_pop$col == "C", "flower.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "A" & 
                                UCD24_pop$row == 17 & 
                                UCD24_pop$col == "C", "flower.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "A" & 
                             UCD24_pop$row == 17 & 
                             UCD24_pop$col == "C", ])

# Directly update the date of the 2023 entry
UCD24_pop[UCD24_pop$bed == "A" & 
                       UCD24_pop$row == 17 & 
                       UCD24_pop$col == "C", "fruit.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "A" & 
                                UCD24_pop$row == 17 & 
                                UCD24_pop$col == "C", "fruit.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "A" & 
                             UCD24_pop$row == 17 & 
                             UCD24_pop$col == "C", ])

```

```{r}
# Directly update the date of the 2014 entry
UCD24_pop[UCD24_pop$bed == "B" & 
                       UCD24_pop$row == 14 & 
                       UCD24_pop$col == "A", "fruit.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "B" & 
                                UCD24_pop$row == 14 & 
                                UCD24_pop$col == "A", "fruit.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "B" & 
                             UCD24_pop$row == 14 & 
                             UCD24_pop$col == "A", ])
```

```{r}
# Directly update the date of the 2014 entry
UCD24_pop[UCD24_pop$bed == "D" & 
                       UCD24_pop$row == 44 & 
                       UCD24_pop$col == "D", "last.FL.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "D" & 
                                UCD24_pop$row == 44 & 
                                UCD24_pop$col == "D", "last.FL.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "D" & 
                             UCD24_pop$row == 44 & 
                             UCD24_pop$col == "D", ])
```

```{r}
#CHECK THIS ONE, LAST FR DATE BEFORE LAST FLOWER DATE!!!!
# Directly update the date of the 2014 entry
UCD24_pop[UCD24_pop$bed == "F" & 
                       UCD24_pop$row == 33 & 
                       UCD24_pop$col == "C", "last.FR.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "F" & 
                                UCD24_pop$row == 33 & 
                                UCD24_pop$col == "C", "last.FR.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "F" & 
                             UCD24_pop$row == 33 & 
                             UCD24_pop$col == "C", ])
```

```{r}
# Directly update the date of the 2014 entry
UCD24_pop[UCD24_pop$bed == "C" & 
                       UCD24_pop$row == 31 & 
                       UCD24_pop$col == "C", "last.FR.date"] <- 
  update(UCD24_pop[UCD24_pop$bed == "C" & 
                                UCD24_pop$row == 31 & 
                                UCD24_pop$col == "C", "last.FR.date"][[1]], 
         year = 2024)

# Verify the updated date
print(UCD24_pop[UCD24_pop$bed == "C" & 
                             UCD24_pop$row == 31 & 
                             UCD24_pop$col == "C", ])
```

```{r}
UCD24_pop %>%
  filter(!is.na(pop.id)) %>%
  ggplot(aes(x=bud.date, fill=pop.id)) +
  geom_histogram()

UCD24_pop %>%
  filter(!is.na(pop.id)) %>%
  ggplot(aes(x=flower.date, fill=pop.id)) +
  geom_histogram()

UCD24_pop %>%
  filter(!is.na(pop.id)) %>%
  ggplot(aes(x=fruit.date, fill=pop.id)) +
  geom_histogram()

UCD24_pop %>%
  filter(!is.na(pop.id)) %>%
  ggplot(aes(x=last.FL.date, fill=pop.id)) +
  geom_histogram()

UCD24_pop %>%
  filter(!is.na(pop.id)) %>%
  ggplot(aes(x=last.FR.date, fill=pop.id)) +
  geom_histogram()
```

### WL2 2023

```{r}
WL2_pheno <- read_csv("../input/WL2_2023_Data/WL2_mort_pheno_20231020_corrected.csv")
```

#### Apply lubridate

```{r}
#apply lubridate
WL2_pheno$bud.date <- mdy(WL2_pheno$bud.date)
WL2_pheno$flower.date <- mdy(WL2_pheno$flower.date)
WL2_pheno$fruit.date <- mdy(WL2_pheno$fruit.date)
#WL2_pheno$last.flower.date <- mdy(WL2_pheno$last.flower.date)
#WL2_pheno$last.fruit.date <- mdy(WL2_pheno$last.fruit.date)
```

#### Filter to only include individuals that budded

```{r}
WL2_pheno_filter <- WL2_pheno[!is.na(WL2_pheno$bud.date),]
```

#### One individual has the wrong fruit date year, correct to say 2023

```{r}
#one of the fruit dates says 2013
fruit2014 <- WL2_pheno_filter[year(WL2_pheno_filter$fruit.date) == 2013,]
print(fruit2014)

# Directly update the date of the 2013 entry
WL2_pheno_filter[WL2_pheno_filter$block == "A" & 
                       WL2_pheno_filter$bed == "A" & 
                       WL2_pheno_filter$bed.row == 23 & WL2_pheno_filter$bed.col == "D", "fruit.date"] <- 
  update(WL2_pheno_filter[WL2_pheno_filter$block == "A" & 
                       WL2_pheno_filter$bed == "A" & 
                       WL2_pheno_filter$bed.row == 23 & WL2_pheno_filter$bed.col == "D", "fruit.date"][[1]], year = 2023)

# Verify the updated date
print(WL2_pheno_filter[WL2_pheno_filter$block == "A" & 
                       WL2_pheno_filter$bed == "A" & 
                       WL2_pheno_filter$bed.row == 23 & WL2_pheno_filter$bed.col == "D", ])
```

#### Visualize bud, flower, and fruit dates

```{r}
ggplot(WL2_pheno_filter, aes(x=bud.date, fill=pop)) +
  geom_histogram()

ggplot(WL2_pheno_filter, aes(x=flower.date, fill=pop)) +
  geom_histogram()

ggplot(WL2_pheno_filter, aes(x=fruit.date, fill=pop)) +
  geom_histogram()
```

### WL2 2024

```{r}
WL2_24 <- read_csv("../input/WL2_2024_Data/WL2_mort_pheno_20241023_corrected.csv")

```

```{r}
#apply lubridate
WL2_24$bud.date <- mdy(WL2_24$bud.date)
WL2_24$flower.date <- mdy(WL2_24$flower.date)
WL2_24$fruit.date <- mdy(WL2_24$fruit.date)
WL2_24$last.FL.date <- mdy(WL2_24$last.FL.date)
WL2_24$last.FR.date <- mdy(WL2_24$last.FR.date)
WL2_24$death.date <- mdy(WL2_24$death.date)
WL2_24$missing.date <- mdy(WL2_24$missing.date)
```

```{r}
#creating column with only pop info
WL2_24_pop <- WL2_24 %>%
  mutate(pop = case_when(
    grepl("^\\w+_\\d+_\\d+$", unique.ID) ~ sub("^(\\w+)_\\d+_\\d+$", "\\1", unique.ID),
    TRUE ~ unique.ID
  ))
```

```{r}
ggplot(WL2_24_pop, aes(x=bud.date, fill=pop)) +
  geom_histogram()

ggplot(WL2_24_pop, aes(x=flower.date, fill=pop)) +
  geom_histogram()

ggplot(WL2_24_pop, aes(x=fruit.date, fill=pop)) +
  geom_histogram()

ggplot(WL2_24_pop, aes(x=last.FL.date, fill=pop)) +
  geom_histogram()

ggplot(WL2_24_pop, aes(x=last.FR.date, fill=pop)) +
  geom_histogram()
```

Plot as proportions of plants that were planted

Likelihood of flowering given budding
