---
title: "Preliminary Data Exploration"
output: html_document
date: "2024-10-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse)
library(lubridate)
```

### UCD 2022-2023 Transplants

```{r}
UCD_pheno_transplant <- read_csv("../input/UCD_Data/Phenology_survey_transplants_20230508_corrected.csv")
```

#### Renaming columns to remove spaces between words

```{r}
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date First Bud"] <- "Date_First_Bud"
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date First Fruit"] <- "Date_First_Fruit"
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date First Flower"] <- "Date_First_Flower"
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date Last Flower"] <- "Date_Last_FLower"
colnames(UCD_pheno_transplant)[colnames(UCD_pheno_transplant) == "Date Last Fruit"] <- "Date_Last_Fruit"

```

#### Apply lubridate

```{r}
UCD_pheno_transplant$Date_First_Bud <- mdy(UCD_pheno_transplant$Date_First_Bud)
UCD_pheno_transplant$Date_First_Flower <- mdy(UCD_pheno_transplant$Date_First_Flower)
UCD_pheno_transplant$Date_First_Fruit <- mdy(UCD_pheno_transplant$Date_First_Fruit)
#UCD_pheno_transplant$Date_Last_Flower <- mdy(UCD_pheno_transplant$Date_Last_Flower) #no last flower dates
#UCD_pheno_transplant$Date_Last_Fruit <- mdy(UCD_pheno_transplant$Date_Last_Fruit) #no last fruit dates
```

#### Filter to only include individuals that budded

```{r}
UCD_pheno_trans_filter <- UCD_pheno_transplant[!is.na(UCD_pheno_transplant$Date_First_Bud) | !is.na(UCD_pheno_transplant$Date_First_Flower), ]
```

#### Visualize bud date - the year for one individual is wrong

```{r}
ggplot(UCD_pheno_trans_filter, aes(x=Date_First_Bud)) +
  geom_histogram()
```

##### Identify which row has the wrong year 

```{r}
# one first bud date seems to be 2022 which is not correct
bud2022 <- UCD_pheno_trans_filter[year(UCD_pheno_trans_filter$Date_First_Bud) == 2022,]
print(bud2022)
```

##### Update the year to 2023

```{r}
# Directly update the date of the 2022 entry
UCD_pheno_trans_filter[UCD_pheno_trans_filter$block == "L1" & 
                       UCD_pheno_trans_filter$row == 4 & 
                       UCD_pheno_trans_filter$col == "A", "Date_First_Bud"] <- 
  update(UCD_pheno_trans_filter[UCD_pheno_trans_filter$block == "L1" & 
                                UCD_pheno_trans_filter$row == 4 & 
                                UCD_pheno_trans_filter$col == "A", "Date_First_Bud"][[1]], 
         year = 2023)

# Verify the updated date
print(UCD_pheno_trans_filter[UCD_pheno_trans_filter$block == "L1" & 
                             UCD_pheno_trans_filter$row == 4 & 
                             UCD_pheno_trans_filter$col == "A", ])

```

#### Visualize bud, flower, and fruit dates

```{r}
ggplot(UCD_pheno_trans_filter, aes(x=Date_First_Bud, fill=pop)) +
  geom_histogram()

ggplot(UCD_pheno_trans_filter, aes(x=Date_First_Flower, fill=pop))+
  geom_histogram()

ggplot(UCD_pheno_trans_filter, aes(x=Date_First_Fruit, fill=pop)) +
  geom_histogram()
```

### UCD 2022-2023 Germinants

```{r}
UCD_pheno_germ <- read_csv("../input/UCD_Data/Phenology_survey_germination_20230515_corrected.csv")
```

#### Renaming columns to remove spaces between words and apply lubridate

```{r}
#Update column names
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date First Bud"] <- "Date_First_Bud"
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date First Fruit"] <- "Date_First_Fruit"
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date First Flower"] <- "Date_First_Flower"
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date Last Flower"] <- "Date_Last_FLower"
colnames(UCD_pheno_germ)[colnames(UCD_pheno_germ) == "Date Last Fruit"] <- "Date_Last_Fruit"

#apply lubridate
UCD_pheno_germ$Date_First_Bud <- mdy(UCD_pheno_germ$Date_First_Bud)
UCD_pheno_germ$Date_First_Flower <- mdy(UCD_pheno_germ$Date_First_Flower)
UCD_pheno_germ$Date_First_Fruit <- mdy(UCD_pheno_germ$Date_First_Fruit)
#UCD_pheno_germ$Date_Last_Flower <- mdy(UCD_pheno_germ$Date_Last_Flower) #no last flower dates
UCD_pheno_germ$Date_Last_Fruit <- mdy(UCD_pheno_germ$Date_Last_Fruit)
```

#### Visualize bud, flower, and fruit dates

```{r}
ggplot(UCD_pheno_germ, aes(x=Date_First_Bud, fill=pop)) +
  geom_histogram()

ggplot(UCD_pheno_germ, aes(x=Date_First_Flower, fill=pop)) +
  geom_histogram()

ggplot(UCD_pheno_germ, aes(x=Date_First_Fruit, fill=pop)) +
  geom_histogram()
```

### WL2 2023

```{r}
WL2_pheno <- read_csv("../input/WL2_Data/WL2_mort_pheno_20231020_corrected.csv")
```

#### Apply lubridate

```{r}
#apply lubridate
WL2_pheno$bud.date <- mdy(WL2_pheno$bud.date)
WL2_pheno$flower.date <- mdy(WL2_pheno$flower.date)
WL2_pheno$fruit.date <- mdy(WL2_pheno$fruit.date)
#WL2_pheno$last.flower.date <- mdy(WL2_pheno$last.flower.date)
#WL2_pheno$last.fruit.date <- mdy(WL2_pheno$last.fruit.date)
```

#### Filter to only include individuals that budded

```{r}
WL2_pheno_filter <- WL2_pheno[!is.na(WL2_pheno$bud.date),]
```

#### One individual has the wrong fruit date year, correct to say 2023

```{r}
#one of the fruit dates says 2013
fruit2014 <- WL2_pheno_filter[year(WL2_pheno_filter$fruit.date) == 2013,]
print(fruit2014)

# Directly update the date of the 2013 entry
WL2_pheno_filter[WL2_pheno_filter$block == "A" & 
                       WL2_pheno_filter$bed == "A" & 
                       WL2_pheno_filter$bed.row == 23 & WL2_pheno_filter$bed.col == "D", "fruit.date"] <- 
  update(WL2_pheno_filter[WL2_pheno_filter$block == "A" & 
                       WL2_pheno_filter$bed == "A" & 
                       WL2_pheno_filter$bed.row == 23 & WL2_pheno_filter$bed.col == "D", "fruit.date"][[1]], year = 2023)

# Verify the updated date
print(WL2_pheno_filter[WL2_pheno_filter$block == "A" & 
                       WL2_pheno_filter$bed == "A" & 
                       WL2_pheno_filter$bed.row == 23 & WL2_pheno_filter$bed.col == "D", ])
```

#### Visualize bud, flower, and fruit dates

```{r}
ggplot(WL2_pheno_filter, aes(x=bud.date, fill=pop)) +
  geom_histogram()

ggplot(WL2_pheno_filter, aes(x=flower.date, fill=pop)) +
  geom_histogram()

ggplot(WL2_pheno_filter, aes(x=fruit.date, fill=pop)) +
  geom_histogram()
```
