---
title: "UCD_Phenology"
output: html_document
date: "2025-02-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse)
library(lubridate)
library(lme4)
library(googlesheets4)
library(lmtest)
library(car)
```

```{r}
UCD22_23_transplant <- read_csv("../input/UCD22_23_transplant.csv")
UCD22_23_transplant_prop <- read_csv("../input/UCD22_23_transplant_prop.csv")
UCD22_23_germ <- read_csv("../input/UCD22_23_germ.csv")
UCD22_23_germ_prop <- read_csv("../input/UCD22_23_germ_prop.csv")

UCD23_24 <- read_csv("../input/UCD23_24.csv")
UCD23_24_prop <- read_csv("../input/UCD23_24_prop.csv")
```

#### Convert Date to Julian Date

```{r}
#UCD 22-23 transplants
UCD22_23_transplant$bud.julian <- as.numeric(format(UCD22_23_transplant$bud.date, "%j"))
UCD22_23_transplant$flower.julian <- as.numeric(format(UCD22_23_transplant$flower.date, "%j"))
UCD22_23_transplant$fruit.julian <- as.numeric(format(UCD22_23_transplant$fruit.date, "%j"))

#UCD 22-23 germinants
UCD22_23_germ$bud.julian <- as.numeric(format(UCD22_23_germ$bud.date, "%j"))
UCD22_23_germ$flower.julian <- as.numeric(format(UCD22_23_germ$flower.date, "%j"))
UCD22_23_germ$fruit.julian <- as.numeric(format(UCD22_23_germ$fruit.date, "%j"))

#UCD 23-24
#filter UCD23_24 for only parent pops
UCD23_24_parent <- UCD23_24 %>%
  filter(!is.na(pop))

UCD23_24_parent$bud.julian <- as.numeric(format(UCD23_24_parent$bud.date, "%j"))
UCD23_24_parent$flower.julian <- as.numeric(format(UCD23_24_parent$flower.date, "%j"))
UCD23_24_parent$fruit.julian <- as.numeric(format(UCD23_24_parent$fruit.date, "%j"))
```

#### Plotting Julian day means of onset of phenophases

```{r}
#UCD 22-23 transplant
UCD22_23t_julian <- UCD22_23_transplant %>%
  select(pop, bud.julian, flower.julian, fruit.julian) %>%
  filter(rowSums(is.na(select(., bud.julian, flower.julian, fruit.julian))) < 3) %>%
  gather(key = "phenophase", value = "julian_day", -pop)

ggplot(UCD22_23t_julian, aes(x=julian_day, y=phenophase, fill=phenophase)) +
  geom_boxplot()+
  facet_wrap(~pop)

#UCD 22-23 germinants
UCD22_23g_julian <- UCD22_23_germ %>%
  select(pop, bud.julian, flower.julian, fruit.julian) %>%
  filter(rowSums(is.na(select(., bud.julian, flower.julian, fruit.julian))) < 3) %>%
  gather(key = "phenophase", value = "julian_day", -pop)

ggplot(UCD22_23g_julian, aes(x=julian_day, y=phenophase, fill=phenophase)) +
  geom_boxplot()+
  facet_wrap(~pop)

#UCD 23-24 
UCD23_24_julian <- UCD23_24_parent %>%
  select(pop, bud.julian, flower.julian, fruit.julian) %>%
  gather(key="phenophase", value="julian_day", -pop)

UCD23_24_julian$phenophase <- factor(UCD23_24_julian$phenophase,
                                           levels = c("bud.julian", "flower.julian", "fruit.julian"))

UCD23_24_julian %>%
  filter(!(pop %in% c("LV1", "WL1", "WV", "YO11")))%>%
  ggplot(aes(x=julian_day, y=phenophase, fill=phenophase)) +
  geom_boxplot()+
  facet_wrap(~pop)
```

#### Flowering Duration

```{r}
UCD23_24 <- UCD23_24 %>%
  mutate(FL_duration = as.numeric(last.FL.date_dfp - flower.date_dfp))
```

#### Fruit Duration

```{r}
UCD23_24 <- UCD23_24 %>%
  mutate(FR_duration = as.numeric(last.FR.date_dfp - fruit.date_dfp))
```

#### Total Reproductive Duration

```{r}
UCD23_24 <- UCD23_24 %>%
  mutate(Total_repro_duration = as.numeric(last.FR.date_dfp - bud.date_dfp))


#NEED TO LOOK INTO CALCULATION FOR THIS SINCE SOME TOTAL REPRO LENGTHS ARE SHORTER THAN FLOWERING OR FRUITING
#WHAT TO DO IF PLANT DIES, USE DEATH DATE TO CALCULATE LAST REPRO?
```

```{r}
UCD23_24 %>%
  filter(pop != "NA") %>%
  ggplot(aes(x=pop, y=FL_duration)) +
  geom_boxplot()

UCD23_24 %>%
  filter(pop != "NA") %>%
  ggplot(aes(x=pop, y=FR_duration)) +
  geom_boxplot()

UCD23_24 %>%
  filter(pop != "NA") %>%
  ggplot(aes(x=pop, y=Total_repro_duration)) +
  geom_boxplot()
```

#### Initiate repro by pop for UCD 22-23 Transplant

#### Add climate distance data

```{r}
clim_dist_UCD22_23 <- read_csv("../input/growthseason_GowersEnvtalDist_UCD.csv")
```

```{r}
#change pop column of clim_dist_UCD to match UCD24_pop
colnames(clim_dist_UCD22_23)[colnames(clim_dist_UCD22_23) == "parent.pop"] <- "pop"
                                 
#join clim_dist_UCD to phenology tibble and add binary for initiating reproduction
UCD22_23_model <- UCD22_23_transplant %>%
  filter(pop != "buffer") %>%
  mutate(initiate_repro = if_else(!is.na(bud.date_dfp), 1, 0)) %>%
  left_join(clim_dist_UCD22_23, by="pop")
```

```{r}
m1 <- glm(initiate_repro ~ -1 + pop,
    data= UCD22_23_model,
    family = binomial())
summary(m1)

null1 <- glm(initiate_repro ~ 1,
    data= UCD22_23_model,
    family = binomial())

lrtest(null1, m1)
#pop is a  significant predictor of initiating reproduction, but mostly because only some pops did but many didn't

#plogis(coef(m1))
```

```{r}
#climate distance as a fixed effect and pop as a random
mm1 <- glmer(initiate_repro ~ Gowers_Dist + (1 | pop),
            data=UCD_24_model,
            family=binomial)
summary(mm1)

Anova(mm1)
```

#### 

Schmid et al 2017

Onset of phenophases- date when the respective stage occurred for the first time

Duration of phenophases- number of days between the onset and the day when the stage became dominant (budding?, flowering = last.FL.date - flower.date, fruiting = last.FR.date - fruit.date) WHAT HAPPENS IF IT INITIATED FL OR FR BUT DIED OR NEVER FINISHED - USE DEATH DATE or...

Total duration- number of days between budding and last fruit WHAT HAPPENS IF THERE IS NO LAST_FR_DATE? need to look at last census for last FR date or use death date as end of repro

```{r}


```

ONLY USE TRANSPLANT DATA FROM UCD22-23
